mouse level quality plots

### 2020-01-09 doc/Effects.Rmd
small updates for effects docs

I ran lm of the data from the plots above and found that the coefficient is 10x higher in males, p value max-out low and in females the quality co-efficicent is much smaller and p.value larger.

@@

Each point is a single mouse

@@

The mouse 22jun15_G_m2 has strange pattern, investigate the Python SC outputs.

Some of the pattern goes away when redo.crop is set to equal 'no'.

@@

When this project was started the main mouse room for vendor purchased mice was Biotech. The mouse room changed in summer 2017. The Gough mice were always housed in Charmony. Of the final data set there are three strains had mice housed in different rooms; WSB, LEW, and MSM. (not balanced/ limited comparisons / )

@@

The brackets in the plots above reflect the output of 'compare_means()' which uses a wilcox.test.

I double checked the wilcox p values for wilcox.test and compare_means, they're similar.

WSB single wilcox test pvalue is  Wlcx.ROOM_WSB.female$p.value` which is the same as the raw from compare_means(). The compare_means table shows that WSB is multiplied by 2.

The values look a little strange to me, I don't understand the p.adj values yet, (p values multiplied by the number of comparisons). To me it seems like there are 5 t.tests, . I don't know why there are only 3 values of adjusted p-values)

@@

I think these results indicate that there is not a strong room effect. Also considering that

@@

 - Qualitatively, there is unlikely to be a maternal age effect, but these data are underpowered.

@@

right now the only independent variable is month dissected. Should I enter info about PFA? time in humid chamber?

>>>>

These results confirm that quality has a significant effect on MLH1 count per cell. How this affects further inferences is a little unclear since the quality score is a very rough indicator of technical error.

@@

@@

The mouse 22jun15_G_m2 has strange pattern, investigate the Python SC outputs. Some of the pattern goes away when redo.crop is set to equal 'no'.

@@

When this project was started the main mouse room for vendor purchased mice was Biotech. The mouse room changed in summer 2017. The Gough mice were always housed in Charmony. Of the final data set there are three strains had mice housed in different rooms; WSB, LEW, and MSM. (This data might not balanced and there are limited comparisons.)

@@

The brackets in the plots above reflect the output of 'compare_means()' which uses a wilcox.test. I double checked the wilcox p values for wilcox.test and compare_means, they're similar.

@@

@@

 - Qualitatively, it doesn't appear that there is a maternal age effect, but these data are underpowered.

@@

Plotting euth date and mean CO number.

### 2020-01-09 doc/Main_Report_Results.rmd
small updates for effects docs

---
title: "MLH1 Data Report"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
bibliography: refs_Main_Report.bib
---

@@

#load main data file

load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #
#load(file="data/MetaData.RData") #this was replacing most updated data file

@@

load in more Data, chrm proportions, total SC,
write / standardize the data set up for those

Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
-ene endings refer to the cells/units within the cell stage? (I think)

@@

@@

Details of dataset cleaning ect is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd

@@

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )

@@

figure 1, change to all white, add the subspecies lines below.

```{r first.boxplots, echo=FALSE, fig.align='center'}

@@

#these plot currently look bad
ss
justMuscu
```

@@

*cast female data came from another publication

add code for the table table. may have to deal with the excess mouse issue

@@

These are the basic Mixed Models we used as a frame work to analyze all the MLH1 data.

@@

Seting up the data set for using the mixed model (base data set, HQ dataset)

@@

- male ages within range

Currently I ran through 3 datasets: 1) base, 2) HQ and 3) Molf (from HQ)

@@

Above are the total number of mice in the data set

FOR FINAL ANALYSIS, MOL Will be coded as a seperate subsp (change all of the main model reports to)
For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.

subspecies levels: 2(3), Musc and Domesticus (Molossinus)

strain levels: Dom: 3, Musc: 4  (Mol: 2)

@@

Above is the table of coefficients for MM.1, HQ dataset, lme(), random effects strain and strain specific sex effect.

@@

Above is the table of coefficients for MM.1.mol, molssinus coded as seperate subspecies, lme(), random effects strain and strain specific sex effect.

@@

Above is the table of coefficients for MM.2, HQ dataset, lme(), random effects coded without sex specific effect.
This is model has sex and interaction coefficient is significant.

@@

Above is the table of coefficients for MM.2.2, HQ dataset, gls(), No random effects.  
I think this gls has sex and interaction term as significant.

@@

Above is the table of coefficients for MM.3, HQ dataset, lm function, sex, strain and subsp are interacting fixed effects.

@@

Above is the table of coefficients for **MM.4**, HQ dataset, lm function, just sex and strain, no subspecies.

Significant factors are 
strainG (p=0.0005), strainMSM and strainPWD (very slight),
sex:G (p=0.001),
sex:MSM, Sex:PWD (very slight p=.05, p=0.3),

@@

#### Summary of mean COs

@@

# Variance in CO count

in addition to the mean gwRR, the variance is also an important metric to dicribe

@@

@@

The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect pvalues were larger and only significant for one of the models.

When Molossinus is coded as a seperate subspecies, then interaction terms are no longer significant.

@@

list out the points whic hsuport rapid male specific evolution for MLH1 rates.

@@

chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured")+ylim(c(0,1))

@@

# Sexual dimorphism and heterochiasmy things

total sc length / SC compaction and telomere bias in positioning are conserved / consistantly sexually dimorphism (in these mice?)  for most species

@@

# ""

>>>>

---
title: "MLH1 Results Outline"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document: default
bibliography: refs_Main_Report.bib
---

@@

#load main data file
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_1.5.20.RData")#added batch18

#this MLH1 only goes up to batch16!
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#batch18
#load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/MLH1/MLH1_data_setup_12.20.19.RData")#fixing bugs, formating
#load(file="data/MLH1/MLH1_data_setup_12.3.19.RData")#removed exclude mice
#load(file="data/MLH1/MLH1_data_setup_11.12.19.RData")#added batch17
#load(file="data/MLH1/MLH1_data_setup_11.11.19.RData")#deleted DUP image
#load(file="data/MLH1/MLH1_data_setup_9.26.19.RData")#added batch16
#load(file="data/MLH1/MLH1_data_setup_8.29.19.RData") #

#load(file="data/MetaData.RData") #this was replacing most updated data file

@@

@@

#Terminology Note, leptonema, zygonema, and pachynema diplonema diakinesis, are the stage names.
#-ene endings refer to the cells/units within the cell stage? (I think)

@@

Details of data set cleaning ect is in SetupData.rmd
Effects from different factors are reviewed in Effects.Rmd

@@

cor.test(AP_mouse_table.HQ.10plus$mean_co[AP_mouse_table.HQ.10plus$sex == "female"], 
    AP_mouse_table.HQ.10plus$var[AP_mouse_table.HQ.10plus$sex == "female"] )
```

@@

```{r first.boxplots, echo=FALSE, fig.align='center', warning=FALSE}

@@

#these plot currently look bad
```

@@

*cast female data came from another publication, add this ref (Kohler?)

The bottom lines and subspecies points don't exactly match up.

**add in cell images like the other draft figures.**

@@

These are the basic Mixed Models we used as a frame work to analyze all the MLH1 count data.

@@

Setting up the data set for using the mixed model (base data set, HQ data set)

@@

- male ages within good range (5 - 12? weeks)

Currently I ran through 3 data sets: 1) base, 2) HQ and 3) Molf (from HQ)

@@

FOR FINAL ANALYSIS, make sure that MOL is coded as a separate subsp (change all of the main model reports to). For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species.

subspecies levels: 2(3), Musc and Domestics (Mousiness)

strain levels: Don: 3, Musc: 4  (Mol: 2)

@@

Above is the table of coefficients for MM.1, HQ data set, lme(), random effects strain and strain specific sex effect.

@@

Above is the table of coefficients for MM.1.mol, molssinus coded as separate subspecies, lme(), random effects strain and strain specific sex effect.

@@

Above is the table of coefficients for MM.2, HQ data set, lme(), random effects coded without sex specific effect.
This is model has sex and interaction coefficient is significant.

@@

Above is the table of coefficients for MM.2.2, HQ data set, gls(), No random effects.  
I think this gls has sex and interaction term as significant.

@@

Above is the table of coefficients for MM.3, HQ data set, lm function, sex, strain and subsp are interacting fixed effects.

@@

Above is the table of coefficients for **MM.4**, HQ data set, lm function, just sex and strain, no subspecies.

Significant factors are 
strain G (p=0.0005), strainMSM and strainPWD (very slight),
sex:G (p=0.001),
sex:MSM, Sex:PWD (very slight p=.05, p=0.3),

@@

 Summary of mean COs

@@

## Variance in CO count

in addition to the mean gwRR, the variance is also an important metric to describe

@@

### Figures

@@

The general pattern across all the models, is that sex is a significant effect for all models. The sex-subsp interaction effect p values were larger and only significant for one of the models.

When Molossinus is coded as a separate subspecies, then interaction terms are no longer significant.

@@

list out the points which support rapid male specific evolution for MLH1 rates.

@@

chrm.class.barplot <- ggplot(data = melt.Chrm.Prop.table, aes(y = value, x = strain, fill=as.factor(type)) ) + geom_col(position = "dodge")+facet_grid(sex~strain, scales = "free")+ggtitle("Chrm Proportions from Hand Measured Data")+ylim(c(0,1))

@@

# Sexual dimorphism and heterochiasmy

total SC length / SC compaction and telomere bias in positioning are conserved / consistently sexually dimorphism (in these mice?)  for most species

@@

# Blank

### 2020-01-09 doc/MouseLevelPlots.Rmd
small updates for effects docs

exclude_mice <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                  "3sep18_AST_m1",
                  "3jel18_TOM_m1","3sep18_TOM_m1",
                  "31jul17_HMI_m1",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )

>>>>

exclude_mice_list <- c("10mar15_WSB_m2", "8oct14_WSB_f1", "8oct14_WSB_f3", "18nov17_WSB_f5",
                  "8jun15_G_f3", "8jun15_G_f4",
                  "4jan17_LEW_f2",
                  "8oct14_PWD_f3", "1apr15_PWD_f2", "8oct14_PWD_f5",
                  "12sep16_MSM_f3","31aug16_MSM_m2",
                  "27oct18_MOLF_m3",
                  "26oct18_SKIVE_f1","26oct18_SKIVE_f2","26oct18_SKIVE_f3",
                  "27aug18_CZECH_m1","8jun18_CZECH_m1",
                  "3sep18_AST_m1",
                  "3jel18_TOM_m1","3sep18_TOM_m1",
                  "31jul17_HMI_m1","3sep18_HMI_m1","3sep18_HMI_m2",
                 "21aug17_SPI_f1", "30may18_SPIC_f1", "30oct17_SPIC_m1", "8jun18_SPIC_m2"
                  )
